\textbf{ Interpretable Gaussian Processes for Stellar Variability }

### 2021-01-19 tex/ms.tex
new figure, as if we needed more

\begin{abstract}
    The use of Gaussian processes (GPs) in stellar light curve modeling has recently
    become almost ubiquitous, given their ease of use and flexibility.
    GPs excel in particular at marginalization
    of the stellar signal in cases where the stellar variability is treated
    as a nuisance, such as in exoplanet transit modeling.
    However, in cases where the stellar signal is of primary interest,
    GPs have thus far been less useful, as it is often difficult to
    relate their hyperparameters to the physical properties of interest, such
    as the size, contrast, and latitudinal distribution of star spots.
    Instead, it is common practice to explicitly model the effect
    of individual star spots on the light curve and attempt to infer their
    properties via optimization or posterior inference, a process that is
    ill-posed, plagued by degeneracies, and computationally intractable when
    applied to stars with more than a few spots and/or to ensembles of many
    light curves.
    %
    In this paper, we derive a closed-form expression for the
    mean and covariance of a Gaussian process model that describes
    the light curve of a rotating star conditioned on a given distribution of
    star spot sizes, contrasts, and latitudes.
    %
    We show that our GP is unbiased,
    allowing one to correctly infer physical parameters of interest from one
    or more stellar light curves, including
    the typical radii and the mean and variance of the latitude
    distribution of star spots.
    %
    Our GP has far-ranging implications for understanding the variability and
    magnetic activity of stars from both light curves and radial velocity (RV)
    measurements, as well as for modeling correlated noise in both transiting
    and RV exoplanet searches.
    %
    Our implementation is efficient, user-friendly, and open source, available
    as the \Python package \starryprocess.
    \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
\end{abstract}

@@

Over the past decade, Gaussian processes (GPs) have gained traction as
a leading tool for modeling correlated signals in astrophysical timeseries.
Despite whatever mystique the words ``Gaussian process'' may evoke, a GP
is nothing but a Gaussian distribution in many (formally infinite)
dimensions. Specifically, it is a Gaussian distribution over
\emph{functions} spanning a continuous domain (in our case, the time domain).
Similar to a multivariate Gaussian, which is described by
a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
the mean of the process and a $(K \times K)$
matrix $\pmb{\Sigma}$ describing its covariance,
a GP is fully specified by a mean
function $m(t)$ and a kernel function $k(t, t')$.
To say that a random vector-valued variable $\mathbbb{f}$
defined on a $(K \times 1)$ time array $\mathbf{t}$
is ``distributed as a GP'' means that we may write
%
\begin{align}
    \label{eq:fnormal}
    \mathbbb{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
    \quad,
\end{align}
%
where the elements of the mean and covariance are given by
$\mu_i = m(t_i)$ and $\Sigma_{i,j} = k(t_i, t_j)$, respectively.%
\footnote{%
    In this paper, we will use blackboard font (i.e., $\mathbbb{f}$) to
    denote random variables and serif font (i.e., $\mathbf{f}$) to denote
    particular realizations of those variables.
    See Appendix~\ref{sec:notation} for a detailed explanation of our notation.
}
Because of this relationship to multivariate Gaussians,
GPs are extremely easy to sample from.%
\footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
    sampling from the corresponding GP (if it exists)
    can be done in a single line of code by calling
    \texttt{numpy.random.multivariate\_normal(mean, cov)}.}
But the real showstopper is the application of GPs to inference problems.
Multivariate Gaussian distributions have a closed-form (marginal) likelihood
function, so it is easy (and extremely efficient) to compute the probability
of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
(i.e., the ``likelihood''; see Equation~\ref{eq:log-like} below).
This can in turn be maximized
to infer the optimal values of the model parameters
or used in a
numerical sampling scheme to compute the probability of those parameters
given the data (i.e., the ``posterior'').
Thanks to modern computer architectures, linear algebra packages, and
GP algorithms,
evaluating the GP likelihood may typically be done in a fraction of a second
for a reasonably-sized dataset (i.e., $K \lesssim 10^4$ datapoints).

@@

Which brings us, finally, to an important drawback of GPs when applied
to modeling stellar variability. While it is straightforward to derive
posterior constraints on the hyperparameters of a GP, it is usually
not obvious what those constraints actually tell us about the star
or the actual source of the variability. While the GP period $P$
may sometimes be interpreted as the rotational period of the star,%
\footnote{An exception to this is in the presence of strong differential
    rotation, in which case many periods may be present in the data, or
    when spots evolve coherently, which can also introduce weak periodicities
    in the light curve.}
the relationship between quantities like the amplitude $\sigma^2$
and the timescale $\tau$ and physical quantites of interest---such
as the properties of the star spots---is much fuzzier.
It may be tempting to interpret $\sigma^2$ as some measure of the spot
contrast or the total number of spots, or $\tau$ as the timescale on which the
spots evolve, but there are no guarantees these interpretations will
hold in general. After all, the choice of kernel is quite often \emph{ad hoc},
providing an \emph{effective}---as opposed to \emph{interpretable}---description
of the physics.

For this reason, GPs are usually employed to model stellar variability
when the variability itself is a nuisance parameter. For example, if the
goal is to constrain properties of a transiting exoplanet, a GP might be
used to remove (or, better yet, to marginalize over) the stellar signal.
In this case, the physics behind the variability is irrelevant, so an
effective model of this sort is generally sufficient. When the goal is to understand
the variability itself, however, GPs have been far less useful to date. Instead,
the usual approach has been to explicitly model the stellar signal: in the
case of star spot-induced variability, one must construct a physical model
for the stellar surface and for how it projects into the light curve as the
star rotates. This might entail directly modeling the star spots,
with free parameters controling the spot size, contrast, position,
and temporal evolution of each of the $n$ spots, where $n$ is also
a free parameter. In addition to this being a numerically challenging
problem for all but the smallest values of $n$, the problem is
intrinsically ill-posed, meaning there are strong degeneracies at play
that formally preclude a unique solution to the stellar map.
After all, a stellar light curve is a one-dimensional representation of a
two-dimensional (or, if the star evolves in time, three-dimensional)
surface, and thus it simply cannot encode all the information about the stellar surface.
In fact, the \emph{vast majority} of the modes on the stellar surface
are in the \emph{null space}, the set of surface features that have identically
zero effect on the light curve.
The underdetermined nature of the stellar mapping problem is therefore fundamental: it
is simply not possible to uniquely determine the properties of all the spots on
a star from its rotational light curve.

However, it is hardly ever the case that this is actually our end goal.
After all, physics predicts the properties of stellar surfaces at a fairly
high level: i.e., typical spot sizes, active spot latitudes, or approximate
timescales on which spots evolve. We are hardly ever interested in the
\emph{particular} properties of a \emph{particular} spot, as we wouldn't really
know what to do with that information! Instead, we often treat
(whether explicitly or not)
the properties of a star spot as a draw from some parent distribution
controlling (say) the average and spread in the radii of the spots.
The parameters controling this distribution are the ones that we can
predict with physics; they are therefore also the ones we are usually
interested in.

@@

Before we dive in, it is useful to introduce the spherical harmonics,
a set of orthogonal functions on the surface of the sphere which we will use
to describe the intensity field on the surface of a star
(Figure~\ref{fig:ylms}). As we will see below,
the spherical harmonics are a particularly convenient basis in which to
describe star spot distributions%
\footnote{There are, of course, drawbacks to using this basis:
    in particular, the spherical harmonics are smooth, continuous functions that
    struggle (at finite degree $l$) to capture high resolution features such as
    small star spots. We discuss this point at length in \S\ref{sec:tinyspots},
    where we show that our model is useful even when applied to stars with spots
    smaller than the effective resolution of the GP.}, as they will allow us to compute
moments of the intensity distribution analytically. Of more immediate concern,
\citet{Luger2019} showed that there is a linear relationship between the
spherical harmonic expansion of a stellar surface and the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe as the star rotates about a fixed axis.
If the stellar surface intensity is described by the spherical harmonic
coefficient vector $\mathbf{y}$, the flux is given by
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \mathbf{y}
    \quad,
\end{align}
%
where $\mathbf{1}$ is the ones vector and
$\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is
a function of the stellar inclination $I$, the stellar
rotation period $P$, and the stellar limb darkening coefficients $\mathbf{u}$
(see Appendix~\ref{sec:starry} for details).

@@

are the mean and covariance of the distribution over spherical harmonic coefficients.

@@

The PDFs for the spot radius, the spot contrast, and the number
of spots are delta functions centered at $r$, $c$, and $n$, respectively

@@

the PDF for the latitude $\phi$ of the spots is a Beta distribution in
$\cos\phi$ with (normalized) shape parameters $a$ and $b$,

@@

The second point is of more immediate concern, since we must correct our
expression for the covariance matrix
$\pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)$ if we are to
use our GP to model normalized light curves. For definiteness, let us
consider only mean-normalized light curves.%
\footnote{In practice, the expressions derived below also work well
    for median-normalized light curves, since the distribution of the GP sample median
    is usually close to the distribution of sample mean.}
Since we are dividing the
flux by the mean, one might imagine that we could simply divide the
covariance matrix by the square of the mean and call it a day. However,
this is incorrect, since we must be careful to distinguish between the
\emph{sample} mean and the \emph{process} mean. When normalizing a light
curve by the mean, the operation we perform is
%
\begin{align}
    \label{eq:ftilde}
    \tilde{\mathbbb{f}\hspace{0.2em}} & = \frac{\mathbbb{f}}{\left<\mathbb{f}\right>}
    \quad,
\end{align}
%
where $\tilde{\mathbbb{f}\hspace{0.2em}}$ is the normalized, unit-mean light curve,
$\mathbbb{f}$ is the measured light curve (in detector counts), and
$\left<\mathbb{f}\right>$ is the \emph{sample} mean: i.e., the average value of
a given star's light curve (which we model as a sample from our GP).
This may be close to but is in general different from the \emph{process} mean,
$\pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)$, since the mean of
a draw from the GP is itself normally distributed with a variance that scales
with the GP variance.

@@

What Equation~(\ref{eq:SigmaTilde}) allows us to do is effectively marginalize over
the unknown baseline by modeling the \emph{normalized} flux as a draw
from a Gaussian process:
%
\begin{align}
    \tilde{\mathbbb{f}\hspace{0.2em}}
    \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \sim
    \mathcal{N}\left(
    \mathbf{1},
    \tilde{\pmb{\Sigma}} \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \right)
    \quad.
\end{align}
%
This is appropriate as long as $z \ll 1$, for which the true
distribution of $\tilde{\mathbbb{f}\hspace{0.2em}}$ is approximately Gaussian. In practice,
we recommend employing this trick only for $z \lesssim 0.02$, for which the
error in the approximation to the covariance is less than $10^{-6}$.
In cases where the light curve variability exceeds a few percent, we recommend
modeling the multiplicative amplitude in each light curve as a latent variable, as discussed
above.

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/normgp.pdf}
        \oscaption{normgp}{%
            An example of a flux covariance matrix $\pmb{\Sigma}$ for a
            \starry process (left) and the corresponding covariance of the
            normalized process (right), computed from Equation~(\ref{eq:SigmaTilde}).
            In addition to an offset and an overall scaling relative to the
            original covariance matrix, the covariance of the normalized process
            is discernibly non-stationary.
            \label{fig:normgp}
        }
    \end{centering}
\end{figure}

@@

@@

But before we dive in, it is worth pausing for a moment to take another
look at Figure~\ref{fig:samples}. While we focused on the shortcomings of
the GP as a prior in the discussion above, it is important to appreciate
that it even works in the first place! Despite some issues, the fact that
the surface map samples \emph{actually look like maps of spotted stellar surfaces}
is fairly amazing. A Gaussian process, after all, is a non-parameteric
\emph{process} describing a smooth and continuous function only via its
covariance structure. The GP knows nothing about the existence of discrete
spots---only how any two points on the surface are correlated.
Because spherical harmonics are smooth functions with support over the
entire sphere, the GP also does not know about features restricted
to certain latitudes; in fact, in most applications of GPs to mapping
problems in astronomy \citep[such as in models of the cosmic microwave
    background][]{?}, the process is assumed to be isotropic,
with no preferred angular direction.
The fact that we are able to construct a GP that captures discrete
features restricted to certain regions on the surface (albeit with
some issues) is remarkable.

@@

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/nongaussianity.png}
        \oscaption{nongaussianity}{%
            Corner plot showing the joint distributions of select spherical
            harmonic coefficients corresponding to the $l=15$ expansion of
            $5\times 10^{4}$ stellar surface maps drawn from a certain
            star spot distribution (black points and contours).
            At $l=15$ there are 256 coefficients in total;
            we chose five of the coefficients with the most odd-looking
            distributions to illustrate the non-Gaussianity of the process.
            In addition to non-linear correlations, skewness, and the
            existence of points of very high curvature, some of the distributions
            are also multi-modal.
            The orange contours show slices of the Gaussian approximation to
            the joint distribution; this is the approximation adopted in
            this paper.
            \label{fig:nongaussianity}
        }
    \end{centering}
\end{figure}

It is not immediately obvious that the GP approach should work, because the
true marginal likelihood function
%
$p ( \tilde{\mathbbb{f}\hspace{0.2em}} \, \big| \, P, \mathbf{u}, \pmb{\theta}_\bullet )$
%
is certainly \emph{not} Gaussian. To see why, let us generate $10^{4}$ stellar
surfaces sampled from the true distribution we are trying to model: that is,
a surface with $n=5$ discrete circular spots of fixed, uniform contrast $c=0.1$
and radius $r=20^\circ$ at latitudes $\mu_\phi \pm \sigma_\phi = 30^\circ \pm 5^\circ$.
Let us then expand each surface in spherical harmonics and visualize the
distribution of coefficients $\mathbbb{y}$. Figure~\ref{fig:nongaussianity}
shows the joint distribution for five of the coefficients with the most
non-Gaussian marginal distributions (selected by eye). Different slices
through this distribution
(in black) are skewed, strongly peaked, non-linearly correlated, and even bimodal.
Our approach in this paper is to model this distribution as a Gaussian
(orange contours). While this may be a good approximation in certain regions
of parameter space, it is certainly a poor approximation in others.
In all fairness, the vast majority of the spherical harmonic coefficients
have marginal distributions that \emph{look} to be very Gaussian; however,
those that do not could in principle lead to errors and biases when using
the Gaussian approximation to the likelihood function.

@@

Our sampling parameters are the number of spots $n$, their contrast $c$,
their radius $r$, and the shape parameters $a$ and $b$ describing their
distribution in latitude. As we discuss in Appendix~\ref{sec:lat},
the parameters $a$ and $b$ are easier to sample in than
the mode $\mu_\phi$ and standard deviation $\sigma_\phi$, provided we
account for the Jacobian of the transformation (Equation~\ref{eq:J})
in our log probability function, which maps a uniform prior on $a$ and $b$
to a uniform prior on $\mu_\phi$ and $\sigma_\phi$.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/calibration_default_latitude.pdf}
        \oscaption{calibration_default}{%
            Posterior distributions for the spot latitudes for the default run. Each blue curve corresponds
            to the Beta distribution PDF (Equation~\ref{eq:phi-pdf}) for the spot latitude
            with shape parameters drawn from the posterior in $\mu_\phi$ and $\sigma_\phi$
            (Figure~\ref{fig:calibration_default_corner}); the (hyper)distribution of blue curves
            quantifies our beliefs about how spots are distributed on any given star. The
            orange curve is the true distribution used to generate the spots (see Table~\ref{tab:synthetic}).
            \label{fig:calibration_default_latitude}
        }
    \end{centering}
\end{figure}

>>>>

\begin{abstract}
    The use of Gaussian processes (GPs) in stellar light curve modeling has recently
    become almost ubiquitous, given their ease of use and flexibility.
    GPs excel in particular at marginalization
    of the stellar signal in cases where the
    variability due to star spots rotating in and out of view is treated
    as a nuisance, such as in exoplanet transit modeling.
    However, in cases where the star spot signal is of primary interest,
    GPs have thus far been less useful: it is often difficult to
    relate their hyperparameters to the physical properties of interest, such
    as the size, contrast, and latitudinal distribution of the spots.
    Instead, it is common practice to explicitly model the effect
    of individual star spots on the light curve and attempt to infer their
    properties via optimization or posterior inference. Unfortunately,
    this process is
    degenerate, ill-posed, and often computationally intractable when
    applied to stars with more than a few spots and/or to ensembles of many
    light curves.
    %
    In this paper, we derive a closed-form expression for the
    mean and covariance of a Gaussian process model that describes
    the light curve of a rotating, evolving stellar surface
    conditioned on a given distribution of
    star spot sizes, contrasts, and latitudes.
    %
    We show that our GP is unbiased,
    allowing one to correctly infer physical parameters of interest from one
    or more stellar light curves, including
    the typical radii and the mean and variance of the latitude
    distribution of star spots.
    %
    Our GP has far-ranging implications for understanding the variability and
    magnetic activity of stars from both light curves and radial velocity (RV)
    measurements, as well as for robustly modeling correlated noise in both transiting
    and RV exoplanet searches.
    %
    Our implementation is efficient, user-friendly, and open source, available
    as the \Python package \starryprocess.
    \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
\end{abstract}

@@

Over the past decade, Gaussian processes (GPs) have gained traction as
a leading tool for modeling correlated signals in astrophysical timeseries
\citep[e.g.,][]{RasmussenWilliams2005,BrewerStello2009,Luger2016,ForemanMackey2017,Angus2018}.
Despite whatever mystique the words ``Gaussian process'' may evoke, a GP
is nothing but a Gaussian distribution in many (formally infinite)
dimensions. Specifically, it is a Gaussian distribution over
\emph{functions} spanning a continuous domain (in our case, the time domain).
Similar to a multivariate Gaussian, which is described by
a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
the mean of the process and a $(K \times K)$
matrix $\pmb{\Sigma}$ characterizing its covariance,
a GP is fully specified by a mean
function $m(t)$ and a kernel function $k(t, t')$.
To say that a random vector-valued variable $\mathbbb{f}$
defined on a $(K \times 1)$ time array $\mathbf{t}$
is ``distributed as a GP'' means that we may write
%
\begin{align}
    \label{eq:fnormal}
    \mathbbb{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
    \quad,
\end{align}
%
where the elements of the mean and covariance are given by
$\mu_i = m(t_i)$ and $\Sigma_{i,j} = k(t_i, t_j)$, respectively.%
\footnote{%
    In this paper, we will use blackboard font (i.e., $\mathbbb{f}$) to
    denote random variables and serif font (i.e., $\mathbf{f}$) to denote
    particular realizations of those variables.
    See Appendix~\ref{sec:notation} for a detailed explanation of our notation.
}
Because of this relationship to multivariate Gaussians,
GPs are extremely easy to sample from.%
\footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
    sampling from the corresponding GP (if it exists)
    can be done in a single line of code by calling
    \texttt{numpy.random.multivariate\_normal(mean, cov)}.}
But the real showstopper is the application of GPs to inference problems.
Multivariate Gaussian distributions have a closed-form (marginal) likelihood
function, so it is easy (and extremely efficient) to compute the probability
of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
(i.e., the ``likelihood''; see Equation~\ref{eq:log-like} below).
This can in turn be maximized
to infer the optimal values of the model parameters
or used in a
numerical sampling scheme to compute the probability of those parameters
given the data (i.e., the ``posterior'').
Thanks to modern computer architectures, linear algebra packages, and
GP algorithms,
evaluating the GP likelihood may typically be done in a fraction of a second
for a reasonably-sized dataset (i.e., $K \lesssim 10^4$ datapoints).

@@

Which brings us, finally, to an important drawback of GPs when applied
to modeling stellar spot-induced variability. While it is straightforward to derive
posterior constraints on the hyperparameters of a GP, it is usually
not obvious what those constraints actually tell us about the star
or its surface. While the GP period $P$
may sometimes be interpreted as the rotational period of the star,%
\footnote{An exception to this is in the presence of strong differential
    rotation, in which case many periods may be present in the data, or
    when spots evolve coherently, which can also introduce weak periodicities
    in the light curve.}
the relationship between quantities like the amplitude $\sigma^2$
and the timescale $\tau$ and physical properties of the spots is much fuzzier.
It may be tempting to interpret $\sigma^2$ as some measure of the spot
contrast or the total number of spots, or $\tau$ as the timescale on which the
spots evolve, but there are no guarantees these interpretations will
hold in general. After all, the choice of kernel is quite often \emph{ad hoc},
providing an \emph{effective}---as opposed to \emph{interpretable}---description
of the physics.

For this reason, GPs are usually employed to model stellar variability
when the variability itself is a nuisance parameter. For example, if the
goal is to constrain properties of a transiting exoplanet, a GP might be
used to remove (or, better yet, to marginalize over) the stellar signal.
In this case, the physics behind the variability is irrelevant, so an
effective model of this sort is generally sufficient. When the goal is to understand
the variability itself, however, GPs have been far less useful to date. Instead,
the usual approach has been to explicitly model the stellar signal: in the
case of star spot-induced variability, one must construct a physical model
for the stellar surface and for how it projects into the light curve as the
star rotates. This might entail directly modeling the star spots,
with free parameters controling the spot size, contrast, position,
and temporal evolution of each of the $n$ spots, where $n$ is also
a free parameter \xxx{(CITE ME)}. In addition to this being a numerically challenging
problem for all but the smallest values of $n$, the problem is
intrinsically ill-posed, meaning there are strong degeneracies at play
that formally preclude a unique solution to the stellar map.
After all, a stellar light curve is a one-dimensional representation of a
two-dimensional (or, if the star evolves in time, three-dimensional)
surface, and thus it simply cannot encode all the information about the stellar surface.
In fact, the \emph{vast majority} of the modes on the stellar surface
are in the \emph{null space}, the set of surface features that have identically
zero effect on the light curve \citep[e.g.,][]{Cowan2013,Luger2019}.
The underdetermined nature of the stellar mapping problem is therefore fundamental: it
is simply not possible to uniquely determine the properties of all the spots on
a star from its rotational light curve.

However, it is hardly ever the case that this is actually our end goal.
After all, physics can be used to predict properties of stellar surfaces at a fairly
high level: i.e., typical spot sizes, active spot latitudes, or approximate
timescales on which spots evolve
\citep[e.g.,][]{Schuessler1996,Solanki2006,Cantiello2019}.
We are hardly ever interested in the
\emph{particular} properties of a \emph{particular} spot, as we wouldn't really
know what to do with that information! Instead, we often treat
(whether explicitly or not)
the properties of a star spot as a draw from some parent distribution
controlling (say) the average and spread in the radii of the spots.
The parameters controling this distribution are the ones that we can
predict with physics; they are therefore also the ones we are usually
interested in.

@@

Before we dive in, it is useful to introduce the spherical harmonics,
a set of orthogonal functions on the surface of the sphere which we will use
to describe the intensity field on the surface of a star
(Figure~\ref{fig:ylms}). As we will see below,
the spherical harmonics are a particularly convenient basis in which to
describe star spot distributions%
\footnote{There are, of course, drawbacks to using this basis:
    in particular, the spherical harmonics are smooth, continuous functions that
    struggle (at finite degree $l$) to capture high resolution features such as
    small star spots. We discuss this point at length in \S\ref{sec:tinyspots},
    where we show that our model is useful even when applied to stars with spots
    smaller than the effective resolution of the GP.}, as they will allow us to compute
moments of the intensity distribution analytically. Of more immediate concern,
\citet{Luger2019} showed that there is a linear relationship between the
spherical harmonic expansion of a stellar surface and the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe as the star rotates about a fixed axis.
If the stellar surface intensity is described by a spherical harmonic
coefficient vector $\mathbf{y}$ (up to a certain degree $l_\mathrm{max}$),
the flux is given by
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \mathbf{y}
    \quad,
\end{align}
%
where $\mathbf{1}$ is the ones vector and
$\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is
a function of the stellar inclination $I$, the stellar
rotation period $P$, and the stellar limb darkening coefficients $\mathbf{u}$
(see Appendix~\ref{sec:starry} for details).

@@

are the mean and covariance of the distribution over spherical harmonic coefficient
vectors $\mathbbb{y}$.

@@

For simplicity, the PDFs for the spot radius, the spot contrast, and the number
of spots are chosen to be delta functions centered at $r$, $c$, and $n$, respectively

@@

the PDF for the latitude $\phi$ of the spots is chosen to be a Beta distribution in
$\cos\phi$ with (normalized) parameters $a$ and $b$,

@@

The second point is of more immediate concern, since we must correct our
expression for the covariance matrix
$\pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)$ if we are to
use our GP to model normalized light curves. For definiteness, let us
consider only mean-normalized light curves.%
\footnote{In practice, the expressions derived below also work well
    for median-normalized light curves, since the distribution of the GP sample median
    is usually close to the distribution of the sample mean.}
Since we are dividing the
flux by the mean, one might imagine that we could simply divide the
covariance matrix by the square of the mean. However,
this is incorrect, since we must be careful to distinguish between the
\emph{sample} mean and the \emph{process} mean. When normalizing a light
curve by the mean, the operation we perform is
%
\begin{align}
    \label{eq:ftilde}
    \tilde{\mathbbb{f}\hspace{0.2em}} & = \frac{\mathbbb{f}}{\left<\mathbb{f}\right>}
    \quad,
\end{align}
%
where $\tilde{\mathbbb{f}\hspace{0.2em}}$ is the normalized, unit-mean light curve,
$\mathbbb{f}$ is the measured light curve (in detector counts), and
$\left<\mathbb{f}\right>$ is the \emph{sample} mean: i.e., the average value of
a given star's light curve (which we model as a sample from our GP).
This may be close to but is in general different from the \emph{process} mean,
$\pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)$, since the mean of
a draw from the GP is itself normally distributed with a variance that scales
with the GP variance.%
\footnote{
    Importantly, the sample mean and process mean will be different even in the
    absence of measurement error! In other words, the mean flux of a given
    star (i.e., the sample mean)
    will in general be different from the mean flux \emph{across all stars} with similar
    surface properties (the process mean).
}

@@

What Equation~(\ref{eq:SigmaTilde}) allows us to do is effectively marginalize over
the unknown baseline by modeling the \emph{normalized} flux as a draw
from a Gaussian process:
%
\begin{align}
    \tilde{\mathbbb{f}\hspace{0.2em}}
    \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \sim
    \mathcal{N}\left(
    \mathbf{1},
    \tilde{\pmb{\Sigma}} \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \right)
    \quad.
\end{align}
%
This is appropriate as long as $z \ll 1$, for which the true
distribution of $\tilde{\mathbbb{f}\hspace{0.2em}}$ is approximately Gaussian. In practice,
we recommend employing this trick only for $z \lesssim 0.02$, for which the
error in the approximation to the covariance is less than $10^{-6}$.
In cases where the light curve variability exceeds about ten percent, we recommend
modeling the multiplicative amplitude in each light curve as a latent variable, as discussed
above.

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/normgp.pdf}
        \oscaption{normgp}{%
            An example of a flux covariance matrix $\pmb{\Sigma}$ for a
            \starry process (left) and the corresponding covariance of the
            normalized process (right), computed from Equation~(\ref{eq:SigmaTilde}).
            In addition to an offset and an overall scaling relative to the
            original covariance matrix, the covariance of the normalized process
            is discernibly non-stationary (see Figure~\ref{fig:nonstationarity}).
            \label{fig:normgp}
        }
    \end{centering}
\end{figure}

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/nonstationarity.pdf}
        \oscaption{nonstationarity}{%
            An example of why normalized light curves are non-stationary.
            The top panel shows $1{,}000$ samples from a unit-mean sinusoid with
            an amplitude of 10\% and a period of 10 days, much longer than the
            1 day observation baseline. The bottom panel shows the same light curves,
            each normalized to its own mean. Because the mean tends to be near
            the center of the observation window, points near $t=0.5$ are driven
            to values very close to unity, while points near the edges have much
            larger scatter.
            \label{fig:nonstationarity}
        }
    \end{centering}
\end{figure}

@@

To understand why normalized light curves are non-stationary, it is useful to
consider the
toy example in Figure~\ref{fig:nonstationarity}, in which we draw
$1{,}000$ samples from a sinusoid with period ten times longer than the
observational baseline. In this limit, each light curve is approximately
linear, which causes its mean value to roughly coincide with the midpoint of the
observation window. Division by the mean value results in points near the
midpoint being driven to unity and points near the edges (whose values differ
the most from the mean) to be driven to both large and small values. The
variance across all samples is now distinctly phase-dependent.
%
In general, the non-stationarity will be most extreme in cases where the
observation window is shorter than or comparable to one period, and
asymptotically vanishes in the limit that a large number of cycles are
observed.

@@

But before we dive in, it is worth pausing for a moment to take another
look at Figure~\ref{fig:samples}. While we focused on the shortcomings of
the GP as a prior in the discussion above, it is important to appreciate
that it even works in the first place!
%Despite some issues, the fact that
%the surface map samples \emph{actually look like maps of spotted stellar surfaces}
%is fairly amazing.
A Gaussian process, after all, is a non-parameteric
\emph{process} describing a smooth and continuous function only via its
covariance structure. The GP knows nothing about the existence of discrete
spots---only how any two points on the surface are correlated.
Because spherical harmonics are smooth functions with support over the
entire sphere, the GP also does not know about features restricted
to certain latitudes; in fact, in most applications of GPs to mapping
problems in astronomy \citep[such as in models of the cosmic microwave
    background; e.g.,][]{Wandelt2012}, the process is assumed to be isotropic,
with no preferred angular direction.
The fact that we are able to construct a GP that captures discrete
features restricted to certain regions on the surface (albeit with
some issues) is remarkable.

@@

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/nongaussianity.png}
        \oscaption{nongaussianity}{%
            Corner plot showing the joint distributions of select spherical
            harmonic coefficients corresponding to the $l=15$ expansion of
            $5\times 10^{4}$ stellar surface maps drawn from a certain
            distribution of discrete circular spots (black points and contours).
            At $l=15$ there are 256 coefficients in total;
            we chose five of the coefficients with the most odd-looking
            distributions to illustrate the non-Gaussianity of the process.
            In addition to non-linear correlations, skewness, and the
            existence of points of very high curvature, some of the distributions
            are also multi-modal.
            The orange contours show slices of the Gaussian approximation to
            the joint distribution; this is the approximation adopted in
            this paper.
            \label{fig:nongaussianity}
        }
    \end{centering}
\end{figure}

It is not immediately obvious that the GP approach should work, because the
true marginal likelihood function
%
$p ( \tilde{\mathbbb{f}\hspace{0.2em}} \, \big| \, P, \mathbf{u}, \pmb{\theta}_\bullet )$
%
is certainly \emph{not} Gaussian. To see why, let us generate $10^{4}$ stellar
surfaces sampled from the true distribution we are trying to model: that is,
a surface with $n=5$ discrete circular spots of fixed, uniform contrast $c=0.1$
and radius $r=20^\circ$ at latitudes $\mu_\phi \pm \sigma_\phi = 30^\circ \pm 5^\circ$.
Let us then expand each surface in spherical harmonics and visualize the
distribution of coefficients $\mathbbb{y}$. Figure~\ref{fig:nongaussianity}
shows the joint distribution for five of the coefficients with the most
non-Gaussian marginal distributions (selected by eye). Different slices
through this distribution
(in black) are skewed, strongly peaked, non-linearly correlated, and even bimodal.
Our approach in this paper is to model this distribution as a Gaussian
(orange contours). While this may be a good approximation in certain regions
of parameter space, it is certainly a poor approximation in others.
%
In this section, we will show that, fortunately, the non-Gaussianity of the
distribution is not in general an issue when doing inference with our GP,
as the resulting posteriors are unbiased.

@@

Our sampling parameters are the number of spots $n$, their contrast $c$,
their radius $r$, and the Beta distribution parameters $a$ and $b$ describing their
distribution in latitude. As we discuss in Appendix~\ref{sec:lat},
the parameters $a$ and $b$ are easier to sample in than
the mode $\mu_\phi$ and standard deviation $\sigma_\phi$, provided we
account for the Jacobian of the transformation (Equation~\ref{eq:J})
in our log probability function, which maps a uniform prior on $a$ and $b$
to a uniform prior on $\mu_\phi$ and $\sigma_\phi$.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/calibration_default_latitude.pdf}
        \oscaption{calibration_default}{%
            Posterior distributions for the spot latitudes for the default run. Each blue curve corresponds
            to the Beta distribution PDF (Equation~\ref{eq:phi-pdf}) for the spot latitude
            with parameters drawn from the posterior in $\mu_\phi$ and $\sigma_\phi$
            (Figure~\ref{fig:calibration_default_corner}); the (hyper)distribution of blue curves
            quantifies our beliefs about how spots are distributed on any given star. The
            orange curve is the true distribution used to generate the spots (see Table~\ref{tab:synthetic}).
            \label{fig:calibration_default_latitude}
        }
    \end{centering}
\end{figure}
